---
title: "API Framework with Authentication and GTFS Realtime using NSW Open Data"
author: "Prepared by: kimnewzealand"
date: "Date: 31 October 2017"
output: html_notebook
---

## Synopsis

Whilst working on a real time visualisation project at #ozunconf in Melbourne in October, 2017, we required real time data sources. 

Given a GTFS Realtime data source, the objective of this mini project is to set up a  framework for API data sourcing with GTFS Realtime and authentication with a data visualisation. API documentation varies between APIs so the documentation would need to be reviewed for each API source and the code updated accordingly.

The [httr vignette](https://cran.r-project.org/web/packages/httr/vignettes/api-packages.html) and [Using Web APIs from R](https://www.rstudio.com/resources/videos/using-web-apis-from-r/) have been a very useful references. 


### 

 

* * *

## Setup

**LOAD PACKAGES**

```{r packages,results='markup'}
require("httr")
require("gtfsway")
require(jsonlite)
options(stringsAsFactors = FALSE)
require(data.table)
require(httpuv)
require(leaflet)
require(anytime)
library(htmltools)
```

### 

 

* * *

## Data

[Transport for NSW Open Data](https://opendata.transport.nsw.gov.au/) returns data in GTFS Realtime format.

[GTFS Realtime](https://developers.google.com/transit/gtfs/)  is a feed specification that allows public transportation agencies to provide real time updates about their fleet to application developers. It is an extension to GTFS (General Transit Feed Specification), an open data format for public transportation schedules and associated geographic information. This format is based on [Google protocol buffers](https://developers.google.com/protocol-buffers/).

This API is accessible with an API key requested from the NSW website.

In this example we will be sourcing the ferry locations at a point in time.

### 

 * * *

## Framework 


There are four steps in this framework:

1. Request
2. Parsing the response
3. Data wrangling
4. Visualisation

1 **REQUEST**

**Goal** - For the API request, our Goal is to "Get the real time data from NSW transport".

**Documentation** - We refer to the [Transport for NSW APIs and datasets documentation](https://opendata.transport.nsw.gov.au/documentation) before creating our request.

**API endpoint** - We can identify the API endpoint as https://api.transport.nsw.gov.au/v1/live/.

**Parameters** -

The documentation also states that the API supports 1 parameters:  

debug (Parameter type query) Retrieve protocol messages in a text, text-based format. 
Note return message may be truncated

### 
  
In the request we will use the GET function from the [httr](https://cran.r-project.org/web/packages/httr/index.html) R package, based on the html verb on the endpoint with parameters, to return a response as a list. 

In this request, we will need to  specify the headers for the authentication and the format parameters (accept_json() preferable to using accept_xml() ).

```{r request}
# Set url and parameters
url  <- "https://api.transport.nsw.gov.au/v1/gtfs/vehiclepos/ferries"
APIkey <- "fhHD8fdlKDKJFHKJEOfndkdi89ddjlkjfds9" # This is not the real key
# Construct API request with the url
response <- httr::GET(url = url, httr::accept_json(), httr::add_headers('Authorization'=paste("apikey", APIkey)))
# Look at the content of the list
names(response)
# Confirm that the HTTP status_code is 200 for a successful request
response$status_code
# Check that information is retured as protobuf
http_type(response)
```


2 **PARSING THE RESPONSE**

To parse the content, we will use the gtfs_real time and gtfs_vehiclePosition function  from the [gftsway](https://github.com/SymbolixAU/gtfsway/) R package for parsing the response from RProtoBuf to an R dataset. 

```{r parse}
# Parses the raw response from a GTFS real-time feed
FeedMessage <- gtfs_realtime(response, content = "FeedMessage")
# The function gtfs_tripUpdates() extracts the 'trip_update' feed
ferrylist <- gtfs_vehiclePosition(FeedMessage) 
# Take a look at the results of the first item in the list
ferrylist[[1]]
```


3 **DATA WRANGLING**

Create a dataframe with the nested lists using rbindlist from [data.table](https://cran.r-project.org/web/packages/data.table/index.html) R package.

```{r}
# Create a dataframe 
data <- rbindlist(ferrylist, fill=TRUE)
# Convert the unix epoch time using the anytime package
data$timestamp<- anytime(data$timestamp)
feedtime<- data[1,6]
```

4 **VISUALISATION**


Produce a simple [leaflet](https://rstudio.github.io/leaflet/) plot of the static data.

```{r}
# Create the leaflet widget 
m <- leaflet()  
icons <- awesomeIcons(
  icon = 'ion-android-boat',
  iconColor = 'black',
  library = 'ion'
)
# Then add layers - add map tiles where the default is OpenStreetMap,
m  %>%  
  addTiles(urlTemplate = 'http://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png')  %>%
      addProviderTiles(providers$ Thunderforest.Transport) %>% 
      addAwesomeMarkers(lng = data$lon,lat=data$lat,label = data$vehicle_id,
    labelOptions = labelOptions(noHide = F),icon = icons) %>%
      addLegend(position = 'topright',opacity = 1, 
            colors = 'DODGERBLUE', values=icons,
            labels = 'Sydney Ferry',title=paste("NSW Transport at ",as.character.Date(feedtime)))
      
```

