---
title: "API Basic Framework using ANU Quantum Random Numbers Server"
author: "Prepared by: kimnewzealand"
date: "Date: 31 October 2017"
output: html_notebook
---

## Synopsis

Whilst working on a real time visualisation project at #ozunconf in Melbourne in October, 2017, we required real time data sources and a live random number generator was suggested. 

Given this live streaming data source, the objective of this mini project is to create a basic framework for real time API data sourcing and visualisation. API documentation varies between APIs so each API's documentation would need to be reviewed and the code updated accordingly.

The [httr vignette](https://cran.r-project.org/web/packages/httr/vignettes/api-packages.html) and [Using Web APIs from R](https://www.rstudio.com/resources/videos/using-web-apis-from-r/) have been a very useful references. 



### 

 

* * *

## Setup

**LOAD PACKAGES**

```{r packages}
library(devtools)
require("httr")
require("jsonlite")
```

### 

 

* * *

## Data

The [ANU Quantum Random Numbers Server](https://qrng.anu.edu.au) has an API with live, truly random numbers returned in a JSON format. This API is freely accessible and does not have any API keys.


### 

 * * *

## Framework 


There are four steps in this framework:

1. Request
2. Parsing the response
3. Data wrangling
4. Visualisation  
  
1 **REQUEST**

**Goal** - For the API request, our Goal is to "Get the random numbers from ANU server".

**Documentation** - First we refer to the [ANU Quantum Random Numbers Server API documentation](https://qrng.anu.edu.au/API/api-demo.php) before creating our request.

**API endpoint** - We can identify the API endpoint as https://qrng.anu.edu.au/API/.

**Parameters** - The documentation states that the API supports 3 parameters:  

- _Data type_, the data type must be 'uint8' (returns integers between 0–255), 'uint16' (returns integers between 0–65535) or 'hex16' (returns hexadecimal characters between 00–ff).  
- _Array length_, the length of the array to return. Must be between 1–1024.  
- _Block size_, only needed for 'hex16' data type. Sets the length of each block. Must be between 1–1024.  

In this request, we do not need to not specify the headers or the format parameters (accept_json() and accept_xml() ).

**Method** - In the request we will use the html verb (method) : GET. The GET function from the [httr](https://cran.r-project.org/web/packages/httr/index.html) R package to return a response as a list. 

```{r APIrequest}
# Set ANU Quantum Random Numbers Server url with the length of 30 and type as uint8 parameters
url  <- "https://qrng.anu.edu.au/API/jsonI.php?length=30&type=uint8"
# Construct API request with the url
response <- httr::GET(url = url)
# Look at the content of the list
names(response)
head(response)
# Confirm that the HTTP status_code is 200 for a successful request
response$status_code
# Check that information is retured as JSON
http_type(response)
```

2 **PARSING THE RESPONSE**

To parse the content from JSON, we will use the httr content function and mime type fromJSON from the [jsonlite](https://cran.r-project.org/web/packages/jsonlite/index.html) R package for parsing the response from JSON to an R dataset. 

```{r parse}
# Parse content with as="text"", expecting it to default to UTF-8
parsed <- fromJSON(httr::content(response,"text"))
# Take a look at the parsed list
head(parsed)
```


3 **DATA WRANGLING**

Here we just need to extract the data as a vector from the parsed list.

```{r data}
# Create a vector for the data
data <- parsed$data
data
```

4 **VISUALISATION**


Produce a very simple plot of the static data.

```{r simpleplot}
plot(data)
```



